---
title: "Flu_forecasting 2"
author: "Minh Tam Hoang"
date: "6/25/2019"
output: html_document
---
Note: please embed functions that are required to run the codes below in the document if you want to knit this file.

Flu forecasting results for the analysis of multivariate times series that only include wili_ny_upstate and wili_jfk. VAR and VARMA models are utilized to predict the behaviour of flu in the next four weeks. Since cross-validations for VARMA models take a significant amount of time to return the results, I only comment out which VARMA models are recommended based on Ljung-Box statistics. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include=TRUE}
library(dplyr)
library(forecast)
library(smooth)
library(MTS)
library(urca)
library(tidyverse)
flu_data <- readRDS("data/ny_flu_wiki.rds")
wili_ts1 <- cbind.data.frame(flu_data[,4], flu_data[,5])
wili_flu_ts <- ts(wili_ts1, start = c(2010, 40), frequency = 52)
autoplot(wili_flu_ts)
```
```{r, include=TRUE}
wili_flu_ts%>% diff(lag = 52) -> wili_flu_s
wili_flu_s %>% ur.kpss()%>% summary()
wili_stimulate <- Eccm(wili_flu_s, maxp = 5, maxq = 6)
VARorder(wili_flu_s,maxp = 5)
```
Since seasonality exist in the series with a frequency of 52, seasonal difference is applied first.

Based on results from Unit Root Test to determine, the series now is stationary.
Another differencing is not necessary.

p-values from table of Extended Cross-correlation Matrices suggest a VAR(2), VAR(3), and VAR(4). Other possible choices include VARMA(2,1) and VARMA(1,2).
First, fit a VAR(3) model for seasonally differenced data (lag = 52)

```{r, include=TRUE}
wili_fit1 <- VAR(wili_flu_s, p = 3)

wili_fc_1 <- VARpred(wili_fit1, h = 4)
invert_fc(wili_fc_1$pred, wili_flu_ts, freq = 52, h = 4)->wili_fc1
print("The forecasted values are ")
print(wili_fc1)
CI_function(wili_fc_1$pred, wili_fc_1$se.err,h = 4, ts_frequency = 52, wili_flu_ts, lambda = 37)

wili_CV_rmse1 <- list()
for ( i in 1:4){
  cross_valid(wili_flu_ts, window = 156, h = i, val_p = 3, val_q = 0, freq = 52, lamb = 37)-> wili_cv_rmse
  wili_CV_rmse1[[i]] <- wili_cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(wili_CV_rmse1[[i]]) <- c("wili_jfk", "wili_ny_upstate")
  print(unlist(wili_CV_rmse1[[i]]))
}
```
First, fit a VAR(4) model for seasonally differenced data (lag = 52)
```{r, include=TRUE}
wili_fit2 <- VAR(wili_flu_s, p = 4)

wili_fc_2 <- VARpred(wili_fit2, h = 4)
invert_fc(wili_fc_2$pred, wili_flu_ts, freq = 52, h = 4)->wili_fc2
print("The forecasted values are ")
print(wili_fc2)
CI_function(wili_fc_2$pred, wili_fc_2$se.err,h = 4, ts_frequency = 52, wili_flu_ts, lambda = 37)

wili_CV_rmse2 <- list()
for ( i in 1:4){
  cross_valid(wili_flu_ts, window = 156, h = i, val_p = 4, val_q = 0, freq = 52, lamb = 37)-> wili_cv_rmse
  wili_CV_rmse2[[i]] <- wili_cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(wili_CV_rmse2[[i]]) <- c("wili_jfk", "wili_ny_upstate")
  print(unlist(wili_CV_rmse2[[i]]))
}
```


As the data show the variation that seasonally fluctuates with the seasonality of time series, log transformation is applied right before seasonally difference.< To stablize the variance and linearize the fit>

```{r, include=TRUE}
wili_transformed <- BoxCox(wili_flu_ts, lambda = 0)
wili_transformed %>% diff(lag = 52)-> wili_ts_trans
wili_guess <- Eccm(wili_ts_trans, maxp = 5, maxq = 6)
VARorder(wili_ts_trans,maxp = 5)
```


According to p-values from two-way table, possible choices inlude VAR(3) and VAR(4). VAR(2) and VAR(2,3) may be options, too.
```{r, include=TRUE}
wili_fit3 <- VAR(wili_ts_trans, p = 3)

wili_fc_3 <- VARpred(wili_fit3, h = 4)
invert_fc(wili_fc_3$pred, wili_transformed, freq = 52, h = 4)->wili_fc3
print("The forecasted values are ")
print(InvBoxCox(wili_fc3, lambda = 0))
CI_function(wili_fc_3$pred, wili_fc_3$se.err,h = 4, ts_frequency = 52, wili_flu_ts, lambda = 0)

wili_CV_rmse3 <- list()
for ( i in 1:4){
  cross_valid(wili_flu_ts, window = 156, h = i, val_p = 3, val_q = 0, freq = 52, lamb = 0)-> wili_cv_rmse
  wili_CV_rmse3[[i]] <- wili_cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(wili_CV_rmse3[[i]]) <- c("wili_jfk", "wili_ny_upstate")
  print(unlist(wili_CV_rmse3[[i]]))
}
```
VAR(4):



```{r, include=TRUE}
wili_fit4 <- VAR(wili_ts_trans, p = 4)

wili_fc_4 <- VARpred(wili_fit4, h = 4)
invert_fc(wili_fc_4$pred, wili_transformed, freq = 52, h = 4)->wili_fc4
print("The forecasted values are ")
print(InvBoxCox(wili_fc4, lambda = 0))
CI_function(wili_fc_4$pred, wili_fc_4$se.err,h = 4, ts_frequency = 52, wili_flu_ts, lambda = 0)

wili_CV_rmse4 <- list()
for ( i in 1:4){
  cross_valid(wili_flu_ts, window = 156, h = i, val_p = 4, val_q = 0, freq = 52, lamb = 0)-> wili_cv_rmse
  wili_CV_rmse4[[i]] <- wili_cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(wili_CV_rmse4[[i]]) <- c("wili_jfk", "wili_ny_upstate")
  print(unlist(wili_CV_rmse4[[i]]))
}
```

We might want to try power transformation to transform our data and then apply seasonal differences. Possible choices include VARMA(1,1), VAR(2), VAR(3)
```{r, include=TRUE}
wili_lambda_s <- BoxCox.lambda(wili_flu_ts)
wili_trans_ts <- BoxCox(wili_flu_ts, lambda = wili_lambda_s)
wili_trans_ts %>% diff(lag = 52) %>% ur.kpss() %>% summary()
wili_trans_ts %>% diff(lag = 52) -> wili_trans_ts_diff
w_guess2 <- Eccm(wili_trans_ts_diff, maxp = 5, maxq = 6)
VARorder(wili_trans_ts_diff,maxp = 5)
```

VAR(3):

```{r, include=TRUE}
wili_fit5 <- VAR(wili_trans_ts_diff, p = 3)

wili_fc_5 <- VARpred(wili_fit5, h = 4)
invert_fc(wili_fc_5$pred, wili_trans_ts, freq = 52, h = 4)->wili_fc5
print("The forecasted values are ")
print(InvBoxCox(wili_fc5, lambda = wili_lambda_s))
CI_function(wili_fc_5$pred, wili_fc_5$se.err,h = 4, ts_frequency = 52, wili_flu_ts, lambda = wili_lambda_s)

wili_CV_rmse5 <- list()
for ( i in 1:4){
  cross_valid(wili_flu_ts, window = 156, h = i, val_p = 3, val_q = 0, freq = 52, lamb = wili_lambda_s)-> wili_cv_rmse
  wili_CV_rmse5[[i]] <- wili_cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(wili_CV_rmse5[[i]]) <- c("wili_jfk", "wili_ny_upstate")
  print(unlist(wili_CV_rmse5[[i]]))
}
```

VAR(2)
```{r, include=TRUE}
wili_fit6 <- VAR(wili_trans_ts_diff, p = 2)

wili_fc_6 <- VARpred(wili_fit6, h = 4)
invert_fc(wili_fc_6$pred, wili_trans_ts, freq = 52, h = 4)->wili_fc6
print("The forecasted values are ")
print(InvBoxCox(wili_fc6, lambda = wili_lambda_s))
CI_function(wili_fc_6$pred, wili_fc_6$se.err,h = 4, ts_frequency = 52, wili_flu_ts, lambda = wili_lambda_s)

wili_CV_rmse6 <- list()
for ( i in 1:4){
  cross_valid(wili_flu_ts, window = 156, h = i, val_p = 2, val_q = 0, freq = 52, lamb = wili_lambda_s)-> wili_cv_rmse
  wili_CV_rmse6[[i]] <- wili_cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(wili_CV_rmse6[[i]]) <- c("wili_jfk", "wili_ny_upstate")
  print(unlist(wili_CV_rmse6[[i]]))
}
```


