---
title: "Flu Forecasting"
author: "Minh Tam Hoang"
date: "6/23/2019"
output:
  html_document: default
  pdf_document: default
---
Note: please embed functions that are required to run the codes below in the document if you want to knit this file.

Flu forecasting results for the analysis of multivariate times series that include wiki_cough, wiki_cold, and wiki_influenza. VAR and VARMA models are utilized to predict the behaviour of flu in the next four weeks. Since cross-validations for VARMA models take a significant amount of time to return the results, I only write down the code for VARMA(2,1) for non-transformed data. However, I still comment out which VARMA models are recommended based on Ljung-Box statistics. 
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```



```{r setup, include=TRUE}
library(dplyr)
library(forecast)
library(smooth)
library(MTS)
library(urca)
library(tidyverse)
flu_data <- readRDS("data/ny_flu_wiki.rds")
ts1 <- cbind.data.frame(flu_data[,4], flu_data[,5], flu_data[,6], flu_data[,7], flu_data[,8])
flu_ts <- ts(ts1, start = c(2010, 40), end = c(2015,31), frequency = 52)
autoplot(flu_ts)
```


```{r, include=TRUE}
flu_ts%>% diff(lag = 52) -> flu_s
flu_s %>% ur.kpss()%>% summary()
stimulate <- Eccm(flu_s, maxp = 5, maxq = 6)
VARorder(flu_s,maxp = 5)
```
Since seasonality exist in the series with a frequency of 52, seasonal difference is applied first.

After that, Unit Root Test to determine whether another differencing is preformed.  Since the test statistic is smaller than 1% critical value, the series now is stationary.

p-values from table of Extended Cross-correlation Matrices suggest a VAR(3). Other possible choices include VARMA(2,1) and VARMA(1,2).


First, fit a VAR(3) model for seasonally differenced data (lag = 52)
```{r, include=TRUE}
fit1 <- VAR(flu_s, p = 3)

fc_1 <- VARpred(fit1, h = 4)
invert_fc(fc_1$pred, flu_ts, freq = 52, h = 4)->fc1
print("The forecasted values are ")
print(fc1)
CI_function(fc_1$pred, fc_1$se.err,h = 4, ts_frequency = 52, flu_ts, lambda = 37)

CV_rmse1 <- list()
for ( i in 1:4){
  cross_valid(flu_ts, window = 156, h = i, val_p = 3, val_q = 0, freq = 52, lamb = 37)->cv_rmse
  CV_rmse1[[i]] <- cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(CV_rmse1[[i]]) <- c("wili_jfk", "wili_ny_upstate", "wiki_cold", "wiki_cough", "wiki_influenza")
  print(unlist(CV_rmse1[[i]]))
}

```

Now, try a VARMA(2,1) model for seasonally differenced data (lag = 52)
```{r, include = FALSE}
fit2 <- VARMA(flu_s, p = 2, q = 1)


fc_2 <- VARMApred(fit2, h = 4)
invert_fc(fc_2$pred, flu_ts, freq = 52, h = 4)->fc2
print("The forecasted values are ")
print(fc2)
CI_function(fc_2$pred, fc_2$se.err,h = 4, ts_frequency = 52, flu_ts, lambda = 37)
  
  

CV_rmse2 <- list()
for ( i in 1:4){
  cross_valid(flu_ts, window = 156, h = i, val_p = 2, val_q = 1, freq = 52, lamb = 37)->cv_rmse
  CV_rmse2[[i]] <- cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(CV_rmse2[[i]]) <- c("wili_jfk", "wili_ny_upstate", "wiki_cold", "wiki_cough", "wiki_influenza")
  print(unlist(CV_rmse2[[i]]))
}
```

Although the data show variations that do not have any clear trend, transformation may be still useful in linearizing the data and equalizing the variances. First, I apply log transformation

```{r, include=TRUE}
transformed <- BoxCox(flu_ts, lambda = 0)
transformed %>% diff(lag = 52)-> ts_trans
guess <- Eccm(ts_trans, maxp = 5, maxq = 6)

```
As the data have a strong seasonal pattern, seasonal difference of lag 52 is done first. The test statistic is smaller than 2.5% crit value, therefore no differencing is needed. From p-values from two-way table, it is clear that VAR(2) and VAR(3) is recommended. VARMA(1,1) and VARMA(2,1) is also a possible choice.

Fit VAR(2) first
```{r, include= TRUE}
trans_fit1 <- VAR(ts_trans, p = 2)## VAR(2)



transfc_3 <- VARpred(trans_fit1, h = 4)
invert_fc(transfc_3$pred, transformed, freq = 52, h = 4)->fc_trans
print("The forecasted values are ")
print(InvBoxCox(fc_trans, lambda = 0))
CI_function(transfc_3$pred,transfc_3$se.err,h = 4, ts_frequency = 52, transformed, lambda = 0)
  
  

CV_rmse4 <- list()
for ( i in 1:4){
  cross_valid(flu_ts, window = 156, h = i, val_p = 2, val_q = 0, freq = 52, lamb = 0)->cv_rmse
  CV_rmse4[[i]] <- cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(CV_rmse4[[i]]) <- c("wili_jfk", "wili_ny_upstate", "wiki_cold", "wiki_cough", "wiki_influenza")
  print(unlist(CV_rmse4[[i]]))
}


```
Then, try VAR(3)
```{r, include=TRUE}
trans_fit2 <- VAR(ts_trans, p = 3)## VAR(3)


transfc_4 <- VARpred(trans_fit2, h = 4)
invert_fc(transfc_4$pred, transformed, freq = 52, h = 4)->fc_trans_2
print("The forecasted values are ")
print(InvBoxCox(fc_trans_2, lambda = 0))
CI_function(transfc_4$pred,transfc_4$se.err,h = 4, ts_frequency = 52, transformed, lambda = 0)


CV_rmse5 <- list()
for ( i in 1:4){
  cross_valid(flu_ts, window = 156, h = i, val_p = 3, val_q = 0, freq = 52, lamb = 0)->cv_rmse
  CV_rmse5[[i]] <- cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(CV_rmse5[[i]]) <- c("wili_jfk", "wili_ny_upstate", "wiki_cold", "wiki_cough", "wiki_influenza")
  print(unlist(CV_rmse5[[i]]))
}


```

We might want to try power transformation to transform our data and then apply seasonal differences. Possible choices include VARMA(1,1), VAR(2), VAR(3)
```{r, include=TRUE}
lambda_s <- BoxCox.lambda(flu_ts)
trans_ts <- BoxCox(flu_ts, lambda = lambda_s)
trans_ts %>% diff(lag = 52) %>% ur.kpss() %>% summary()
trans_ts %>% diff(lag = 52) -> trans_ts_diff
guess2 <- Eccm(trans_ts_diff, maxp = 5, maxq = 6)
```
Try VAR(3)
```{r, include=TRUE}
trans_fit3 <- VAR(trans_ts_diff, p = 3)## 


transfc_5 <- VARpred(trans_fit3, h = 4)
invert_fc(transfc_5$pred, trans_ts, freq = 52, h = 4)->fc_t3
print("The forecasted values are ")
print(InvBoxCox(fc_t3, lambda = lambda_s))
CI_function(transfc_5$pred,transfc_5$se.err,h = 4, ts_frequency = 52, trans_ts, lambda = lambda_s)


CV_rmse6 <- list()
for ( i in 1:4){
  cross_valid(flu_ts, window = 156, h = i, val_p = 3, val_q = 0, freq = 52, lamb = lambda_s)->cv_rmse
  CV_rmse6[[i]] <- cv_rmse
}
for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(CV_rmse6[[i]]) <- c("wili_jfk", "wili_ny_upstate", "wiki_cold", "wiki_cough", "wiki_influenza")
  print(unlist(CV_rmse6[[i]]))
}


```

Try VAR(2)
```{r, include=TRUE}
trans_fit4 <- VAR(trans_ts_diff, p = 2)## 


transfc_6 <- VARpred(trans_fit4, h = 4)
invert_fc(transfc_6$pred, trans_ts, freq = 52, h = 4)-> fc4
print("The forecasted values are ")
print(InvBoxCox(fc4, lambda = lambda_s))
CI_function(transfc_6$pred,transfc_6$se.err,h = 4, ts_frequency = 52, trans_ts, lambda = lambda_s)



CV_rmse7 <- list()
for ( i in 1:4){
  cross_valid(flu_ts, window = 156, h = i, val_p = 2, val_q = 0, freq = 52, lamb = lambda_s)->cv_rmse
  CV_rmse7[[i]] <- cv_rmse
}

for (i in 1:4){
  print("Cross-validated RMSE for forecast horizon ")
  print(i)
  names(CV_rmse7[[i]]) <- c("wili_jfk", "wili_ny_upstate", "wiki_cold", "wiki_cough", "wiki_influenza")
  print(unlist(CV_rmse7[[i]]))
}


```
